OA Share

### 2021-01-25 analysis/manuscript.Rmd
kableextra with conditional color :-)

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
```

@@

This section first presents the results of our analysis of hybrid OA uptake in Elsevier's journal portfolio with a view to licensing, disciplinary differences, and citation impact. Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different sponsor types, institutional sectors, and individual funders.

@@

### Journal-level analysis{.unnumbered}

@@

Figure \ref{fig:number_of_els_journals} shows that in Elsevier's OA portfolio of fully OA and hybrid OA journals, the latter has outweighed the former consistently throughout the five years, with only little variation overall. In 2015, Elsevier published `r els_jns_count %>% filter(year == 2015, oa_model == "Open Access") %>% pull(n)` fully OA journals and `r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(n)` hybrid journals (`r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(prop)*100`%). Since then, Elsevier's hybrid journals increased in number by `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n)) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% pull(n))` journals (n=`r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` in 2020) and slightly decreased in proportion to Elsevier's overall OA journal portfolio by 1% (`r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(prop), 2) * 100`%; `r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` out of `r els_jns_count %>% filter(year == 2020) %>% pull(n) %>% sum()` journals).

@@

We identified `r filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` OA journals that reverse flipped--transitioned from fully OA to hybrid [@Matthias_2019]. We excluded these journals from the following article-level analysis; if we had not, OA articles published in an OA journal before it reverse-flipped would have been falsely attributed to the hybrid model. Likewise, we identified and excluded journals flipping from hybrid to fully OA in the same timeframe (n=`r filter(flipped_jns_ind, oa_model == "Open Access") %>% pull(n)`). Our final sample for the article-level analysis consisted of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` unique hybrid journals.

### Article-level analysis{.unnumbered}

@@

Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article. Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, representing `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`). Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period between 2015-19, illustrating a moderate growth in hybrid OA over time. The number of hybrid journals with at least one OA article increased each year, as did the number of OA articles, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` 10,672 in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019. However, since the total number of articles published in these journals also increased over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.

```{r overview}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~` `,
    "Elsevier Hybrid Journals with ≥ 1 OA article", "n_journals", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_n", "Total ",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_sd", "SD",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_n", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_sd", "SD",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake", "Total (%)",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_mean", "Avg.",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind, -ind_group) %>%
  kableExtra::kbl(booktabs = T, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                  caption = "Overview of hybrid open access uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
                  position = "H"
) %>%
  kable_styling() %>%
  pack_rows("Elsevier Hybrid Journals with OA article", 1, 1) %>%
  pack_rows("Articles", 2, 4) %>%
  pack_rows("OA Articles", 5, 7) %>%
  pack_rows("OA Uptake", 8, 10)
```

@@

Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year, showing considerable variation among Elsevier's hybrid journals. The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots). Notably, upper quartiles and whiskers stretch farther over the years, indicating that the range of uptake among journals with an above-average proportion of OA articles increased over the years. Nevertheless, the prevalence of hybrid OA is relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.

@@

### Share of the total volume published in OA{.unnumbered}

@@

During the same period, Elsevier provided OA to `r all_oa` articles, corresponding to a share of `r all_oa / all_els * 100` of the total article volume (n=`r all_els`). The OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind fully OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r length(unique(mirror_jn$container_title))` mirror journals that published `r sum(mirror_jn$all_articles)` articles, and Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) both in number and share of OA articles (see Figure \ref{fig:oa_volume_fig}-A).

```{r oa_volume_fig, fig.cap="Open Access in Elsevier journals in terms of article volume between 2015-2019. Figure A presents the share of open access (OA) articles relative to Elsevier's overall article output for each OA type from 2015-2019 and by open content license. Figure B presents the aggregated OA article volume of Elsevier journals by OA type."}
# see analysis/006_license.Rmd how the figure was created

@@

### License prevalence{.unnumbered}

@@

As far as we observed, Elsevier licensed OA articles in hybrid journals under two possible Creative Commons (CC) licenses—CC-BY or CC-BY-NC-ND. CC-BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited. In contrast, CC-BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives. In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC-BY-NC-ND license marginally increased and maintained a greater share than that of CC-BY (see Figure \ref{fig:oa_volume_fig}-B). Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid OA had the highest number and proportion of openly available articles licensed under CC-BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to fully OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program, Open Archive (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) . Most of these delayed OA articles were provided under an Elsevier user license (Els-User), prohibiting reuse for commercial purposes.

### Hybrid OA uptake across subjects{.unnumbered}

@@

Next, we present disciplinary variations in hybrid OA. Table \ref{tab:subject_area_table} presents high-level findings by AJSC subject area. In the five-years period studied, the Physical Science (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and Health Sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article published.  In terms of article volume, most OA articles were published in hybrid journals belonging to the Life Sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while Social Sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles. However, it must be noted that there is a large overlap between the journals’ subject areas. Therefore, Table also assigns hybrid journals and OA articles fractionally to a subject area in addition to full counting when a journal belongs to more than one field. In the following, we apply full counting, meaning that journals and articles associated with more than one subject in the AJSC classification were counted once for each area or field.

```{r subject_area_table}
# join
inner_join(subject_area_all_jn, subject_area_all_article, by = "subject_area") %>%
  arrange(subject_area) %>%
  select(1, journals_n, articles_n, journals_frac, articles_frac) %>%
  filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
  kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
      caption = "Subject area overview using full and fractional counting. Table reports the number of Elsevier hybrid journals with at least one OA article, as well as the number of OA articles published in them between 2015-2019. Not shown is one journal coded as Multidisciplinary that published 61 OA articles. Eleven journals with 37 OA articles could not be matched to the Scopus Source Title list used to obtain ASJC codes.") %>%
  add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
  kable_styling() 
```

@@

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall average (Mdn= `r median_all * 100`; excluding one journal coded as Multidisciplinary). The box plot reveals some variation between the subject areas. Notably, with the exception of Environmental Science and Earth and Planetary Sciences, journals from the Physical Sciences show a lower median proportion of hybrid OA articles, whereas most Life Sciences and Social Sciences journals ranked above average. Figure X also shows large variations among the 26 subject fields, with the highest OA rates in Immunology and Microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), Environmental Science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and Neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), and the lowest in Chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), Materials Science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and Dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`).

@@

### Citation impact{.unnumbered}

@@

Furthermore, we were interested in the relationship between a journal's field-specific impact and its OA uptake. Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journal's OA uptake that year. Considering only journals with at least one OA article, we found a weak positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).

@@

### Invoice types

@@

# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100) %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

Table \ref{tab:invoicing_overview} shows that authors (`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) were by far the most frequent invoicing type for hybrid OA, followed by funding bodies/institutional agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%). Interestingly, we also found a small number of cases where hybrid APCs were waived (`r filter(sponsored_art, fct_source == "Fee Waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee Waived") %>% pull(prop)`%).

@@

Figure X illustrates that over the years, Elsevier has increasingly and, to a greater extent, invoiced hybrid OA APCs to authors directly rather than to funders or consortia ("Agreement"). The share of fee-waived articles remained relatively stable at around 6%. However, we found different types of waivers. Around 52% of fee waivers were linked to an OA sponsor, indicating a third-party paid the APC, whereas 48% of Elsevier waived articles did not disclose an OA sponsor. For instance, 856 waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors.  In terms of licensing, Figure X reveals that the type of invoicing was strongly related to the open content license variant. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=32,086; 77%), whereas most articles with agreed payment were licensed under the more permissive CC BY license (n=18,331; 76%).

>>>>

```{r libraries}
# libraries
library(tidyverse)
library(here)
library(cowplot)
library(gt)
library(janitor)
library(kableExtra)
library(stringr)
```

@@

This section first presents the results of our analysis of hybrid OA uptake in Elsevier's journal portfolio with a view to licensing, disciplinary differences, and citation impact. Then, we present a descriptive analysis of Elsevier's invoicing data, highlighting licensing and disciplinary differences for different invoicing channels, institutional sectors, and individual invoice recipients.

@@

### Journal-Level Analysis{.unnumbered}

@@

Figure \ref{fig:number_of_els_journals} visualizes  Elsevier's OA portfolio throughout the five years from 2015-2019, and shows that hybrid journals consistently outweighed fully OA journals, with only little variation overall. In 2015, Elsevier published `r els_jns_count %>% filter(year == 2015, oa_model == "Open Access") %>% pull(n)` fully OA journals and `r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(n)` hybrid journals (`r els_jns_count %>% filter(year == 2015, oa_model == "Hybrid") %>% pull(prop)*100`%). Since then, hybrid journals grew in number by `r (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n)) - (els_jns_count %>% filter(oa_model == "Hybrid", year == 2015) %>% pull(n))` journals (n=`r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` in 2020), resulting in a slightly decreased share of Elsevier's overall OA journal portfolio by 1% (`r round(els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(prop), 2) * 100`%; `r els_jns_count %>% filter(year == 2020, oa_model == "Hybrid") %>% pull(n)` out of `r els_jns_count %>% filter(year == 2020) %>% pull(n) %>% sum()` journals).  Of these, `r filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` journals transitioned from fully OA to hybrid.  For the article-level analysis, we excluded these reverse-flip journals, focussing on `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` unique hybrid journals without OA business model change for the remaining results.

@@

### Article-Level Analysis{.unnumbered}

@@

Between 2015 and 2019, `r hybrid_oa_volume %>% distinct(issn) %>% nrow()` of `r els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n)` (`r  hybrid_oa_volume %>% distinct(issn) %>% nrow() / (els_jns_count %>% filter(oa_model == "Hybrid", year == 2020) %>% pull(n) - filter(flipped_jns_ind, oa_model == "Hybrid") %>% pull(n))*100`%) hybrid journals published at least one OA article. Together these journals published `r five_years_ind %>% pull(oa_n)` OA articles, which represents `r five_years_ind %>% pull(oa_uptake)*100`% of their total article volume (n=`r five_years_ind %>% pull(articles_n)`). Table \ref{tab:overview} presents these findings for each year and the aggregated five-year period, illustrating a moderate growth in hybrid OA over time. With each year, the number of hybrid journals with at least one OA article increased, as did the number of OA articles, which nearly doubled from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_n)` in 2015 to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_n)` in 2019. However, since the total number of articles published in these journals also grew over time, the relative share of OA articles in hybrid journals only increased slightly from `r ind_per_year %>% filter(issued_year == 2015) %>% pull(oa_uptake)*100`% to `r ind_per_year %>% filter(issued_year == 2019) %>% pull(oa_uptake)*100`%.

```{r overview}
table_descriptives_ind <- tribble(
  ~ind_group, ~ind, ~` `,
    "Elsevier Hybrid Journals with ≥ 1 OA article", "n_journals", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_n", "Total ",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "articles_sd", "SD",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_n", "Total",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_mean", "Avg.",
  "Articles in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_sd", "SD",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake", "Total (%)",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_mean", "Avg.",
  "OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA article", "oa_uptake_sd", "SD"
)
inner_join(table_descriptives_ind, ind_summary_table, by = c("ind" = "indicator")) %>%
  dplyr::select(-ind, -ind_group) %>%
  kableExtra::kbl(booktabs = TRUE, format.args = list(big.mark = ","), digits = 1, escape = TRUE,
                  caption = "Overview of hybrid open access uptake across Elsevier hybrid journals with at least one OA article 2015-2019.",
                  position = "H"
) %>%
  kable_styling() %>%
  pack_rows("Elsevier Hybrid Journals with ≥ 1 OA article", 1, 1) %>%
  pack_rows("Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 2, 4) %>%
  pack_rows("OA Articles in Elsevier Hybrid Journals with ≥ 1 OA Article", 5, 7) %>%
  pack_rows("OA Uptake in Elsevier Hybrid Journals with ≥ 1 OA Article", 8, 10)
```

@@

Figure \ref{fig:boxuptake} presents the hybrid OA uptake by year, showing considerable variation among Elsevier's hybrid journals. The box-and-whiskers plot visualizes the median hybrid OA uptake (horizontal line inside the box), the range of the majority of journals (the box), and individual outliers (dots). Notably, the upper quartiles and whiskers stretch farther over the years, which indicates that the range of uptake among journals with an above-average proportion of OA articles increased over the years. Nevertheless, the prevalence of hybrid OA remained relatively low; in 2019, 95% of Elsevier hybrid journals published ≤`r perc_df %>% filter(issued_year == 2019) %>% pull(p_95) * 100`% of their articles OA.

@@

### Share of the Total Volume published in OA{.unnumbered}

@@

During the same period, Elsevier published `r all_oa` OA articles in total (i.e., in hybrid and fully OA journals), representing `r all_oa / all_els * 100` of the total article volume (n=`r all_els`). Looking at Figure \ref{fig:oa_volume_fig}, it is apparant that the OA article volume of hybrid journals (n=`r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_immediate_articles") %>% pull(prop) * 100`%) lagged behind that of Elsevier's Open Archive program (n=`r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(summary_oa_type, type == "hybrid_delayed_articles") %>% pull(prop) * 100`%) and fully OA journals (n=`r filter(summary_oa_type, type == "full_oa") %>% pull(n)`, `r filter(summary_oa_type, type == "full_oa") %>% pull(prop) * 100`%), which included `r sum(mirror_jn$all_articles)` articles published in `r length(unique(mirror_jn$container_title))` mirror journals.

```{r oa_volume_fig, fig.cap="Open Access in Elsevier journals in terms of article volume between 2015-2019. Figure A presents the aggregated OA article volume of Elsevier journals from 2015-2019 by OA type. Figure B presents the share of open access (OA) articles relative to Elsevier's overall article output by OA type and open content license from 2015-2019."}
# see analysis/006_license.Rmd how the figure was created

@@

### License Prevalence{.unnumbered}

@@

As far as we observed, OA articles in Elsevier hybrid journals were published under two possible CC licenses--CC-BY or CC-BY-NC-ND. CC-BY allows others to distribute, remix, adapt, and build upon the licensed work, including for commercial purposes, as long as the original author is credited. In contrast, CC-BY-NC-ND is less permissive as it prohibits commercial reuse and derivatives. In the five-year period from 2015-2019, the proportion of hybrid OA articles published under a CC-BY-NC-ND license marginally increased and maintained a greater share than that of CC-BY (see Figure \ref{fig:oa_volume_fig}-B). Interestingly, Figure \ref{fig:oa_volume_fig}-B also reveals that hybrid journals had the highest number and proportion of openly available articles licensed under CC-BY (n=`r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_immediate_articles") %>% pull(prop) *100`%) compared to fully OA journals  (n=`r filter(cc_by_by_type, type == "full_oa") %>% pull(n)`, `r filter(cc_by_by_type, type == "full_oa") %>% pull(prop) *100`%) and Elsevier's delayed OA program (n=`r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(n)`, `r filter(cc_by_by_type, type == "hybrid_delayed_articles") %>% pull(prop) *100`%) . Most  delayed OA articles were provided under an Elsevier user license (Els-User), which prohibits reuse for commercial purposes.

### Hybrid OA Uptake across Subjects{.unnumbered}

@@

Next, we present disciplinary variations in hybrid OA. Table \ref{tab:subject_area_table} presents the high-level findings by AJSC subject area. In the five-year period we studied, the physical sciences (`r filter(subject_area_all_jn, subject_area == "Physical Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Physical Sciences") %>% pull(articles_n)` OA articles) and health sciences (`r filter(subject_area_all_jn, subject_area == "Health Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Health Sciences") %>% pull(articles_n)` OA articles) were the most prevalent in terms of hybrid journals with at least one OA article.  In terms of article volume, most OA articles were published in hybrid journals belonging to the life sciences (`r filter(subject_area_all_jn, subject_area == "Life Sciences") %>% pull(journals_n)` hybrid journals, `r filter(subject_area_all_article, subject_area == "Life Sciences") %>% pull(articles_n)` OA articles), while social sciences  (`r filter(subject_area_all_jn, subject_area == "Social Sciences") %>% pull(journals_n)`  hybrid journals, `r filter(subject_area_all_article, subject_area == "Social Sciences") %>% pull(articles_n)` articles) journals published the lowest number of OA articles. However, it is important to note that there is a large overlap between the journals’ subject areas. Therefore, Table  \ref{tab:subject_area_table} also assigns hybrid journals and OA articles fractionally to a subject area and compares it to the full counting, where journals assigned more than one subject in the AJSC classification were counted once for each field. For the remaining results, we only present the full counting.

```{r subject_area_table}
# join
inner_join(subject_area_all_jn, subject_area_all_article, by = "subject_area") %>%
  arrange(subject_area) %>%
  select(1, journals_n, articles_n, journals_frac, articles_frac) %>%
  filter(subject_area != "Multidisciplinary", !is.na(subject_area)) %>% 
  mutate(subject_area = stringr::str_to_sentence(subject_area)) %>%
  kbl(booktabs = TRUE, col.names = c("Subject area", "Journals", "OA articles", "Journals", "OA articles"),
      caption = "Subject area overview using full and fractional counting. Table reports the number of Elsevier hybrid journals with at least one OA article, and OA articles published from 2015-2019. One multidisciplinary with 61 OA articles and 11 journals with 37 OA articles that  could not be matched to the Scopus Source Title list used to obtain ASJC codes are not included.",
      format.args = list(big.mark = ",")) %>%
  add_header_above(c(" " = 1, "Full counting" = 2, "Fractional counting" = 2)) %>%
  kable_styling() 
```

@@

Figure \ref{fig:oa_sub_uptake} presents the journals' five-year OA uptake grouped by subject area and field and in comparison to the overall median (Mdn= `r median_all * 100`; excluding one journal coded as multidisciplinary). The box plot reveals some variation between the subject areas. Notably, with the exception of environmental science and earth and planetary sciences, journals from the physical sciences show a lower median proportion of hybrid OA articles, whereas most life sciences and social sciences journals ranked above average. Figure X also shows large variations among the 26 subject fields, with the highest OA rates in immunology and microbiology (Mdn=`r filter(oa_per_subject_rank, description == "Immunology and Microbiology") %>% pull(median_prop)*100`), environmental science (Mdn=`r filter(oa_per_subject_rank, description == "Environmental Science") %>% pull(median_prop)*100`), and neuroscience (Mdn=`r filter(oa_per_subject_rank, description == "Neuroscience") %>% pull(median_prop)*100`), whilst  chemistry (Mdn=`r filter(oa_per_subject_rank, description == "Chemistry") %>% pull(median_prop)*100`), materials science (Mdn=`r filter(oa_per_subject_rank, description == "Materials Science") %>% pull(median_prop)*100`), and dentistry (Mdn=`r filter(oa_per_subject_rank, description == "Dentistry") %>% pull(median_prop)*100`) recorded the lowest OA rate.

@@

### Citation Impact{.unnumbered}

@@

Furthermore, we were interested in the relationship between the journals' field-specific impact and their OA uptake. Spearman's rho correlation coefficient was used to assess the relationship between the 2019 Source Normalized Impact per Paper (SNIP) value calculated by Scopus and the journals' OA uptake that year. Considering only journals with at least one OA article, we found a weak but positive correlation between journal impact and OA uptake ($r_s$=`r as.character(round(cor_test_res$estimate,4))`, $p<0.001$), suggesting that the relationship is not very strong, based on our data (see Figure \ref{fig:sniptest}).

@@

### Invoice Channels{.unnumbered}

@@

# aggregate data
sponsored_articles <- hybrid_volume %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "CC-BY", "CC-BY-NC-ND")) %>%
  count(issued_year, fct_source, cc_norm)
all_articles <- hybrid_volume %>%
  group_by(issued_year) %>%
  summarise(n = n())
sponsored_art <- sponsored_articles %>% 
  group_by(fct_source) %>% 
  summarise(nn = sum(n)) %>% 
  #mutate(cumsum = cumsum(nn)) %>% 
  mutate(prop = nn / sum(nn) * 100) %>%
 # mutate(cumprop = cumsum(prop)) %>%
  adorn_totals("row")
```

As can be seen from Table \ref{tab:invoicing_overview}, hybrid OA APC were most often invoiced to authors (n=`r filter(sponsored_art, fct_source == "Author") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Author") %>% pull(prop)`%) and to a lesser extent part of agreements (n=`r filter(sponsored_art, fct_source == "Agreement") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Agreement") %>% pull(prop)`%). Interestingly, we also found a small number of cases where hybrid APCs were waived (n=`r filter(sponsored_art, fct_source == "Fee waived") %>% pull(nn)`; `r filter(sponsored_art, fct_source == "Fee waived") %>% pull(prop)`%).

@@

```{r els_waived}
els_waived_df <- hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") 
```

Figure \ref{fig:invoiceoverview} illustrates that over the years, Elsevier has increasingly invoiced hybrid OA APCs directly to authors compared to research funders or academic consortia ("Agreement"). The share of fee-waived articles remained relatively stable, but we found different types of waivers. Around `r filter(els_waived_df, !is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)* 100`% of fee waivers were linked to an OA sponsor, which indicates that a third party paid the APC. For instance, `r filter(els_waived_df, oa_sponsor_name == "Académie des Sciences") %>% distinct(doi) %>% nrow()` waivers were linked to the French Académie des Sciences, presumably covering OA publication in its society journals for affiliated authors. The remaining `r filter(els_waived_df, is.na(oa_sponsor_name)) %>% distinct(doi) %>% nrow() / nrow(els_waived_df)*100`% of waived articles did not disclose any OA sponsor.

```{r invoiceoverview, fig.cap="Development of fee-based open access (OA) publishing in Elsevier hybrid journals by invoicing type. Colored bars represent the share of OA articles invoiced to authors, agreements, or waived, grouped by Creative Commons license variant. Grey bars show the total number of hybrid OA articles published in Elsevier journals from 2015-2019."}
# plot
sponsored_articles %>%
  mutate(fct_source = stringr::str_to_title(fct_source)) %>%
  ggplot(aes(x = gsub("20", "'", issued_year), y = n), alpha = 0.8) +
  geom_bar(
    data = all_articles,
    color = "transparent",
    stat = "identity",
    fill = "#b3b3b3a0"
  ) +
  geom_bar(aes(fill = cc_norm), color = "white", stat = "identity", alpha = 0.8, size = 0.25,
           position = position_stack(reverse = T)) +
  facet_wrap( ~ fct_source, nrow = 1) +
  scale_fill_manual(values = 
                      c(`CC-BY` = "#B52141",
                        `CC-BY-NC-ND` = "#0093c7"),
                    name = "CC License Variant") +
  labs(x = "Publication Year", y = "Hybrid OA Articles") + 
  scale_y_continuous( 
                     labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right")
```

Moreover, Figure \ref{fig:invoiceoverview} relates the type of invoicing to the open content license variant. When Elsevier invoiced authors directly, most OA articles in hybrid journals were published under a non-commercial license (n=`r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC-BY-NC-ND") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Author", cc_norm == "CC-BY-NC-ND") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Author") %>% pull(n) %>% sum()* 100`%), whereas most articles billed as part of agreements were licensed under the more permissive CC BY license (n=`r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC-BY") %>% pull(n) %>% sum()`; `r filter(sponsored_articles, fct_source == "Agreement", cc_norm == "CC-BY") %>% pull(n) %>% sum() / filter(sponsored_articles, fct_source == "Agreement") %>% pull(n) %>% sum() * 100`%).

### Invoicing variations across disciplines{.unnumbered}
